The Anthropological Notation Ontology (ANNO): A core ontology for annotating human bones and deriving phenotypes

### 2023-06-06 paper.tex
remove axisplane image
add AnthroWorks3D image
add uncorrected table

% add. options: [seceqn,secthm,crcready]
\documentclass[sw]{iosart2x}
\usepackage{todonotes}
\usepackage{cleveref}
\usepackage{csquotes}
\usepackage{aurl}
%\usepackage{dcolumn}

@@

\section{Description of ANNOdc}\label{sec:core}
% Definition/Beschreibung der Kern-Konzepte/Module und -Relationen mit Beispielen, GFO-Einbettung
\begin{figure}[h]
\includegraphics[width=\textwidth]{img/core.pdf}
\caption{ANNOdc ontology}\label{fig:core}
\end{figure}

@@

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{img/axisplane.pdf}
\caption{Anatomical axes and planes}\label{fig:axisplane}
\end{figure}

>>>>

% add. options: [seceqn,secthm,crcready]
\documentclass[sw]{iosart2x}
\usepackage{todonotes}
\usepackage{cleveref}
\usepackage{csquotes}
\usepackage{aurl}
\usepackage{booktabs}
%\usepackage{dcolumn}

@@

\section{Description and Foundation of ANNOdc}\label{sec:core}
% Definition/Beschreibung der Kern-Konzepte/Module und -Relationen mit Beispielen, GFO-Einbettung
\begin{figure}[h]
\includegraphics[width=\textwidth]{img/core.pdf}
\caption{ANNOdc ontology}\label{fig:core}
\end{figure}

\begin{table}[htbp]
  \centering
  \caption{Table Title}
  \begin{tabular}{llll}
    \toprule
    ANNOdc class                    & Examples                & FMA equivalent / ID                         & GFO superclass    \\
    \midrule
    Anatomical entity               & -                       & -                                          & -                 \\
    Physical anatomical entity      & fma:fma61775            & -                                          & Anatomical structure \\
    material anatomical entity      & fma:fma67165            & Material object                            & Spatial anatomical entity \\
    immaterial anatomical entity    & fma:fma67112            & Space entity                               & Anatomical space  \\
    Space region                    & Anatomical surface      & Median Plane                               & -                 \\
    Anatomical surface              & Surface region          & Anatomical line                             & Mandibular Angle (ppam dexter-go dexter-piam dexter), Sagittal Axis \\
    Anatomical line                 & Line region             & Anatomical point                            & Gonion            \\
    Anatomical point                & Point region            & Anatomical plane                            & Median Plane      \\
    Anatomical plane                & Anatomical axis         & Sagittal Axis                              & -                 \\
    Skeleton                        & -                       & Skeleton fma:fma23875                      & Bone structure    \\
    Bone                            & Mandibula               & Bone organ fma:fma5018                      & -                 \\
    Bone part                       & -                       & Protuberantia mentalis                     & Related to Segment of bone organ fma:fma281808, Zone of bone organ fma:fma10483 \\
    Bone compound                   & Cranium                 & Comparable to union of Skeletal system fma:fma23881 and subdivision of skeletal system fma:fma85544 & - \\
    Tooth structure                 & Dentes Permanentes      & Permanent teeth (fma:fma75152; TA2ID:913)   & Tooth             \\
    Tooth                           & Dens caninus            & Tooth fma:fma12516                          & -                 \\
    Tooth part                      & Cuspis dentis           & fma:fma56481; TA2ID:925                     & -                 \\
    Phenotype                       & Height, Sex             & -                                          & Anatomical Property \\
    Attributive                     & Anatomical relation, e.g., Has\_shape relation & -                              & Attributive       \\
    Anatomical qualitative coordinate & Superior, lateral, medial, posterior & Anatomical Property    & Relator           \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{img/aw3d.png}
\caption{3D model of a Cranium (skull) inserted into a placeholder skeleton in \aw{} depicting examples of various ANNOdc classes.}\label{fig:aw3d}
\end{figure}

@@
